Using Incremental Many-to-One Queries to Build a Fast and Tight Heuristic for A* in Road Networks

### 2022-10-07 paper/ch_potentials.tex
typo fixes

\emph{Contraction Hierarchies} (CH) is a two-phase speedup technique to accelerate shortest path computations on road networks through precomputation.
For a detailed discussion we refer to~\cite{gssv-erlrn-12}.
Here, we only briefly introduce the necessary notation and algorithms used in this paper.
In a preprocessing phase, vertices are ordered totally by ``importance'' where more important vertices should lie on more shortest paths.
Intuitively, vertices on highways are more important than vertices on rural streets.
For CH, such an ordering is obtained heuristically.
The position of a vertex in the order is also denoted as its \emph{rank}.
Vertices of higher rank are informally referred to and visualized as ``higher up'' in the hierarchy.
Therefore, an edge where the tail has a lower rank than the head is an \emph{upward} edge.
Analogously, when the head vertex has the lower rank, the edge is said to go \emph{downward}.
Once such an importance order was obtained, all vertices are contracted successively by ascending importance.
To \emph{contract} a vertex means temporarily removing it from the graph while inserting \emph{shortcut edges} between more important neighbours to preserve shortest distances among them.
The result is an \emph{augmented graph} $G^+$ of original edges and shortcuts with weights $w^+$.
We often refer to the augmented graph split into an upward graph $\gchu = (V, \echu)$ containing only upward edges and a downward graph $\gchd = (V, \echu)$ containing only downward edges.

@@

When looking at total running times for a single many-to-one query, including selection times, Lazy RPHAST delivers very competitive performance and is as fast if not faster than all algorithms evaluated in~\cite{delling_et_al:OASIcs:2011:3266}.
For example, Delling et al. report an average RPHAST running time of 1.97\,ms for selection and query combined for $|B| = 2^{14}$ and 28.52\,ms for $|B| = 2^{20}$ on DIMACS Europe.
Lazy RPHAST (on a CH) takes 0.96\,ms and 10.76\,ms respectively, to compute the same distances.
The CCH variant is slightly slower (1.00\,ms and 14.41\,ms) but still significantly faster than RPHAST.
While the numbers are not perfectly comparable due to different benchmark machines,\footnote{
According to the comparison methodology from~\cite{bdgmpsww-rptn-16} (see \url{https://i11www.iti.kit.edu/~pajor/survey/}), the machine used in~\cite{delling_et_al:OASIcs:2011:3266} (SPA-2) is about 30\% slower than ours (we obtained a score of 33\,159\,ms).
However, these scaling factors have to be interpreted very carefully.
They are obtained from one-to-all Dijkstra searches on continental-sized road networks.
This experiment heavily emphasizes memory bandwidth while neglecting other critical factors such as CPU frequency and cache size and speed, which are likely much more critical for algorithms carefully tuned to exploit data locality.
} we can still conclude that Lazy RPHAST is a valuable extension to RPHAST.
It allows for efficiently handling dynamic source sets and is competitive in the setting where both the target and the source set change between queries.
% This is of course not a fair comparison since the approaches in~\cite{delling_et_al:OASIcs:2011:3266} are geared towards having one expensive selection phase and than many faster queries.
% When computing distances from different sources towards the same target set, RPHAST will outperform Lazy RPHAST.
% However, whenever the target set may change between queries, Lazy RPHAST offers a simple and fast alternative.

>>>>

\emph{Contraction Hierarchies} (CH) is a two-phase speedup technique to accelerate shortest path computations on road networks through precomputation.
For a detailed discussion we refer to~\cite{gssv-erlrn-12}.
Here, we only briefly introduce the necessary notation and algorithms used in this paper.
In a preprocessing phase, vertices are ordered totally by ``importance'' where more important vertices should lie on more shortest paths.
Intuitively, vertices on highways are more important than vertices on rural streets.
For CH, such an ordering is obtained heuristically.
The position of a vertex in the order is also denoted as its \emph{rank}.
Vertices of higher rank are informally referred to and visualized as ``higher up'' in the hierarchy.
Therefore, an edge where the tail has a lower rank than the head is an \emph{upward} edge.
Analogously, when the head vertex has the lower rank, the edge is said to go \emph{downward}.
Once such an importance order was obtained, all vertices are contracted successively by ascending importance.
To \emph{contract} a vertex means temporarily removing it from the graph while inserting \emph{shortcut edges} between more important neighbours to preserve shortest distances among them.
The result is an \emph{augmented graph} $G^+$ of original edges and shortcuts with weights $w^+$.
We often refer to the augmented graph split into an upward graph $\gchu = (V, \echu)$ containing only upward edges and a downward graph $\gchd = (V, \echd)$ containing only downward edges.

@@

When looking at total running times for a single many-to-one query, including selection times, Lazy RPHAST delivers very competitive performance and is as fast if not faster than all algorithms evaluated in~\cite{delling_et_al:OASIcs:2011:3266}.
For example, Delling et al. report an average RPHAST running time of 1.97\,ms for selection and query combined for $|B| = 2^{14}$ and 28.52\,ms for $|B| = 2^{20}$ on DIMACS Europe.
Lazy RPHAST on CH takes 0.96\,ms and 10.76\,ms respectively, to compute the same distances.
The CCH variant is slightly slower (1.00\,ms and 14.41\,ms) but still significantly faster than RPHAST.
While the numbers are not perfectly comparable due to different benchmark machines,\footnote{
According to the comparison methodology from~\cite{bdgmpsww-rptn-16} (see \url{https://i11www.iti.kit.edu/~pajor/survey/}), the machine used in~\cite{delling_et_al:OASIcs:2011:3266} (SPA-2) is about 30\% slower than ours (we obtained a score of 33\,159\,ms).
However, these scaling factors have to be interpreted very carefully.
They are obtained from one-to-all Dijkstra searches on continental-sized road networks.
This experiment heavily emphasizes memory bandwidth while neglecting other critical factors such as CPU frequency and cache size and speed, which are likely much more critical for algorithms carefully tuned to exploit data locality.
} we can still conclude that Lazy RPHAST is a valuable extension to RPHAST.
It allows for efficiently handling dynamic source sets and is competitive in the setting where both the target and the source set change between queries.
% This is of course not a fair comparison since the approaches in~\cite{delling_et_al:OASIcs:2011:3266} are geared towards having one expensive selection phase and than many faster queries.
% When computing distances from different sources towards the same target set, RPHAST will outperform Lazy RPHAST.
% However, whenever the target set may change between queries, Lazy RPHAST offers a simple and fast alternative.
